---
title: 'IMDB ratings: Differences between directors'
author: "Riccardo Persico"
date: "18/10/2020"
output: html_document
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
library(reshape2)
library(patchwork)
library(scales)
```



# Are the mean IMDB rating for Steven Spielberg and Tim Burton are the same or not? Let's see...

## Loading the data

```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)
```

## Summarizing it and plotting it

```{r, intervals_graph}

summary_data <- movies %>% 
  group_by(director) %>% 
  filter(director %in% c("Steven Spielberg", "Tim Burton")) %>% 
  summarise(mean = mean(rating, na.rm = TRUE),
            sd = sd(rating, na.rm = TRUE),
            count= n()) %>%
  mutate(se = sd / sqrt(count),
         lower_ci = mean - qt(1 - (0.05 / 2), count - 1) * se,
         upper_ci = mean + qt(1 - (0.05 / 2), count - 1) * se)

summary_data


ggplot(summary_data, aes(x=mean, y=reorder(director, mean)))+
  geom_point(aes(color = director, size = 3))+
  geom_rect(aes(xmin=max(lower_ci), xmax=min(upper_ci), ymin = -Inf, ymax = +Inf ,alpha = .3))+
  geom_errorbar(aes(xmin = lower_ci , xmax = upper_ci, width = 0.1, color = director, size = 1.5))+
  labs(title = "Do Spielberg and Burton have the same IMDB ratings?", subtitle = "95% confidence intervals overlap", y = "", x = "Mean IMDB Rating")+
  theme_bw()+
   geom_label(aes(y=reorder(director,mean),x=lower_ci,
                   label = round(lower_ci,2)),
                hjust=0.4, vjust=-0.4, fill = NA, label.size = NA)+
   geom_label(aes(y=reorder(director,mean),x=upper_ci,
                   label = round(upper_ci,2)),
                hjust=0.4, vjust=-0.4, fill = NA, label.size = NA)+
  theme(legend.position = "none")+
  geom_text(aes(label=round(mean,2), vjust =-2))
 



```
I have already calculated the confidence intervals for the mean ratings of these two directors and as you can see they overlap. 

## Running the hypotesis tests

```{r, t_test}
#NULL hypothesis : Steven Spielberg's mean IMDB rating = Tim Burton's mean IMDB rating

movies <- movies %>% 
  filter(director %in% c("Steven Spielberg", "Tim Burton")) %>% 
  mutate(steven_tim_val = ifelse(director == "Steven Spielberg",1,0))

#glimpsing the means of Tim Burton's ratings (mean in group 0) and Steven Spielberg's ratings (mean in group 1)

t.test(rating ~ steven_tim_val, data = movies)
  
```
The t-test results show that the mean rating for Tim Burton and Steven Spielberg are different, with Steven's rating being higher than Tim Burton's at 7.57 and 6.93 respectively. Although the confidence intervals overlap, we can see that p-value= 0.01 < 0.05 which confirms that the NULL hypothesis can be rejected.

```{r, hypothesis_test}
#getting Tim and Steven data only
steven_tim <- movies %>% 
  group_by(director) %>% 
  filter(director %in% c("Steven Spielberg", "Tim Burton"))

#assigning numerical values to Steven & Tim
obs_diff <- movies %>%
  specify(rating~director) %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

  
null_dist <- movies %>%
  specify(rating ~ director) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")

```
Therefore, simulating from a null distribution with an infer package and t-test also show that the p-value is below 5%, indicating we can reject the null hypothesis. Tim Burton and Steven Spielberg are 95% unlikely to have the same mean IMDB rating.
